Clustering Constraints on the Relative Sizes of Central and Satellite Galaxies

### 2017-11-09 paper/galsize_paper.tex
edited S 6 and 7

%---------------------------------------------------------------------------------------------------
\begin{figure}
\centering
\includegraphics[width=8cm]{FIGS/rvir_only_lensing_prediction.pdf}
\caption{
{\bf Prediction for $\rhalf-$dependence of galaxy lensing.}
Using the $\rvir-$based model described in \S\ref{subsubsec:rvirmodel}, we make predictions for as-yet-unseen measurements of the $\rhalf-$dependence of galaxy lensing of $\mstar-$complete samples, where ``small" and ``large" subsamples are defined as in \S\ref{subsec:sizedef}. As a generic consequence of satellites being smaller than centrals of the same mass, we predict that future lensing measurements of $\mstar-$complete samples will show $\Delta\Sigma$ of small galaxies to be significantly stronger relative to large galaxies of the same stellar mass. To date, the $\rhalf-$dependence of $\Delta\Sigma$ has only been measured for color-selected samples \citep{charlton_etal17}, in which the trend is reversed and weakened: for both blue and red samples, $\Delta\Sigma$ of large galaxies is (slightly) stronger relative to small galaxies. 
}
\label{fig:lensingprediction}
\end{figure}
%-----------------------------------------------------------------------------------------------------

@@

As discussed in \citet{spindler_wake17}, it is entirely possible that size differences between centrals and satellites of the same mass can be accounted for by mutual covariance with an additional variable such as star formation rate or morphological type \citep[see also][for an explicit demonstration of this scenario]{lilly_carollo16}. In particular, suppose that centrals and satellites of the same mass have different early-type fractions as indicated in \citet{weinmann_etal06}, and that early- and late-type galaxies exhibit universal, but distinct, $\mstar-\rhalf$ relations. In such a case, centrals would be larger than satellites of the same mass, even though {\em early-type centrals} would have the same sizes as {\em early-type satellites}. This scenario could also explain the sign-change and weakening of $\rhalf-$dependent clustering measurements in the present work.

@@

\section{Future Directions for Empirical Modeling of Galaxy Size}
\label{sec:future}

@@

The scope of this work is to serve as a pilot study in which we identify the chief ingredients that can influence the $\rhalf-$dependence of galaxy clustering. While our results suggest that satellite evolution {\em prior} to infall drives the correct relative size of centrals vs. satellites, our implementation cannot be correct in quantitative detail: the clustering ratios of this model shown in Figure \ref{fig:clustering_ratio_upshot} only roughly agree with SDSS. For present purposes, we have chosen to focus on exploring bracketing cases of qualitative ingredients, rather than fine-tuning the model.

@@

On the one hand, this degeneracy with the satellite fraction is unfortunate, because it means galaxy size does not leave a pure and unique signature on $\rhalf-$clustering ratios. However, this can also be viewed as an opportunity to extract tighter constraints on $F_{\rm sat}(\mstar),$ which are sorely needed to discriminate between competing models \citep{watson_conroy13}. Traditional galaxy clustering is already being used to validate and/or fit models of the stellar-to-halo-mass relation \citep[e.g.,][]{leauthaud_etal11,moster_etal13,behroozi13_smhm,lehmann_etal15}. Based on our results, we advocate that empirical models for $\mean{\mstar}{\mhalo}$ be supplemented with additional model ingredients for $\median{\rhalf}{\mstar},$ and that the parameters of the composite model be {\em jointly} constrained by measurements of the stellar mass function, $\mstar-$dependent clustering, and $\rhalf-$dependent clustering.

Figure \ref{fig:censatsizeratios} shows that the satellites in $\mhalo\approx10^{12}\msun$ halos can be as large as $\lesssim40-50\%$ smaller than their central galaxy counterparts. Since the satellite fraction varies quite significantly with morphology and/or color, we have demonstrated that it is entirely plausible that most of the observed environmental trends of galaxy size can be understood simply in terms of central vs. satellite size differences. While it remains to be seen whether this explanation ends up comprising the whole story, one thing is now clear: it is not possible to reliably analyze environmental trends of galaxy $\rhalf$ without properly accounting for satellites. %Because of the large systematics that are common in contemporary group-finding methods \citep{campbell_etal15}, we consider forward modeling methods to be a requisite for continued progress on this topic.

@@

\ben
\item Small galaxies cluster more strongly than large galaxies of the same stellar mass. Differences between the clustering of small and large galaxies increase on small scales $R\lesssim1\mpc,$ and decrease with stellar mass.
\item The most important ingredient influencing this signal is the relative size of central and satellite galaxies. The magnitude, scale-dependence, and $\mstar-$dependence of $\rhalf-$dependent clustering provides strong evidence that satellite galaxies are smaller than central galaxies of the same halo mass.
\item A simple empirical model in which $\rhalf$ is set by halo $\rvir$ at the time of peak halo mass exhibits a clustering signal that is strikingly similar to that seen in SDSS.
\item Models in which $\rhalf$ is regulated by $\mstar,$ rather than $\mhalo,$ are grossly discrepant with the observed clustering signal, even when accounting for satellite mass stripping.
\item Taken together, our findings indicate that satellite-specific processes play a sub-dominant role in setting the relative size of centrals and satellites, which instead appears to be largely predetermined at the time of satellite infall.
\een

Our results can be treated as a boundary condition for more complex and fine-grained models of galaxy size, such as semi-analytic models and hydrodynamical simulations. For convenience, Figure \ref{fig:censatsizeratios} provides a simple summary statistic that can be used as a diagnostic for calibrating alternative modeling efforts. We view the present work as a pilot study that motivates a Bayesian inference program to tightly constrain the galaxy size-halo connection with forward modeling techniques, in direct analogy to the literature on the stellar-to-halo-mass relation. Our publicly available python code provides a simple means for cosmological surveys to generate synthetic galaxy populations with realistic sizes across the cosmic web.

@@

We thank the {\tt Astropy} developers for the package-template \citep{astropy}, and {\tt NumPy} \citep{numpy_ndarray}, {\tt SciPy} \citep{scipy}, IPython, Matplotlib, and GitHub for their extremely useful free software.

This research was supported in part by the National Science Foundation under Grant No. NSF PHY11-25915. Work done at Argonne National Laboratory was supported under the DOE contract DE-AC02-06CH11357.

>>>>

%---------------------------------------------------------------------------------------------------
\begin{figure}
\centering
\includegraphics[width=8cm]{FIGS/rvir_only_lensing_prediction.pdf}
\caption{
{\bf Prediction for $\rhalf-$dependence of galaxy lensing.}
Using the $\rvir-$based model described in \S\ref{subsubsec:rvirmodel}, we make predictions for as-yet-unseen measurements of the $\rhalf-$dependence of galaxy lensing of $\mstar-$complete samples, where ``small" and ``large" subsamples are defined as in \S\ref{subsec:sizedef}. As a generic consequence of satellites being smaller than centrals of the same mass, we predict that future lensing measurements of $\mstar-$complete samples will show $\Delta\Sigma$ of small galaxies to be significantly stronger relative to large galaxies of the same stellar mass. To date, the $\rhalf-$dependence of $\Delta\Sigma$ has only been measured for color-selected samples \citep{charlton_etal17}, in which the trend is reversed and weakened: for both blue and red samples, $\Delta\Sigma$ of large galaxies is (slightly) stronger relative to small galaxies. \textcolor{blue}{AK: is it worth showing a second panel with $\Delta\Sigma$ for color-selected samples to back up the last sentence?}
}
\label{fig:lensingprediction}
\end{figure}
%-----------------------------------------------------------------------------------------------------

@@

\textcolor{blue}{AK: when I was reading this paragraph, I had trouble digesting its meaning... Also it definitely broke the flow of the narrative for me.}
As discussed in \citet{spindler_wake17}, it is entirely possible that size differences between centrals and satellites of the same mass can be accounted for by mutual covariance with an additional variable such as star formation rate or morphological type \citep[see also][for an explicit demonstration of this scenario]{lilly_carollo16}. In particular, suppose that centrals and satellites of the same mass have different early-type fractions as indicated in \citet{weinmann_etal06}, and that early- and late-type galaxies exhibit universal, but distinct, $\mstar-\rhalf$ relations. In such a case, centrals would be larger than satellites of the same mass, even though {\em early-type centrals} would have the same sizes as {\em early-type satellites}. This scenario could also explain the sign-change and weakening of $\rhalf-$dependent clustering measurements in the present work.

@@

%-----------------------------------------------------------------------------------------
\section{Future Directions for Empirical Modeling of Galaxy Size}
\label{sec:future}
%-----------------------------------------------------------------------------------------

@@

The scope of this work is to serve as a pilot study in which we identify the chief ingredients that can influence the $\rhalf-$dependence of galaxy clustering. While our results suggest that sizes of galaxies are linearly correlated with the
virial radius of halos at the moment the halo mass reached its peak and that sizes of satellite galaxies were set {\em prior} to infall, our model is probably does not capture galaxy properties and their trends in full quantitative detail. 
For example, the clustering ratios of this model shown in Figure \ref{fig:clustering_ratio_upshot} do capture the overall trends, but also don't match the observed statistics fully. 
 For present purposes, we have chosen to focus on exploring bracketing cases of qualitative ingredients, rather than fine-tuning the model.

@@

On the one hand, this degeneracy with the satellite fraction is unfortunate, because it means galaxy size does not leave a pure and unique signature on $\rhalf-$clustering ratios. However, this can also be viewed as an opportunity to extract tighter constraints on $F_{\rm sat}(\mstar),$ which are sorely needed to discriminate between competing models \citep{watson_conroy13}. Traditional galaxy clustering is already being used to validate and/or fit models of the stellar-to-halo-mass relation \citep[e.g.,][]{leauthaud_etal11,moster_etal13,behroozi13_smhm,lehmann_etal15}. Based on our results, we advocate that empirical models for $\mean{\mstar}{\mhalo}$ be supplemented with additional model ingredients for $\median{\rhalf}{\mstar},$ (\textcolor{blue}{AK:} did you mean $\langle\rhalf|\mhalo\rangle$ here? Seems more logical in the context of this sentence) and that the parameters of the composite model be {\em jointly} constrained by measurements of the stellar mass function, $\mstar-$dependent clustering, and $\rhalf-$dependent clustering.

Figure \ref{fig:censatsizeratios} shows that the satellites in $\mhalo\approx10^{12}\msun$ halos can be as large as $\lesssim40-50\%$ smaller than their central galaxy counterparts. Since the satellite fraction varies quite significantly with morphology and/or color, we have demonstrated that it is entirely plausible that most of the observed environmental trends of galaxy size can be understood simply in terms of central vs. satellite size differences. The latter reflects not evolution of galaxies inside host halos, but is due to the fact that satellites have reached their peak mass at higher redshifts than the centrals. While it remains to be seen whether this explanation ends up comprising the whole story, one thing is now clear: it is not possible to reliably analyze environmental trends of galaxy $\rhalf$ without properly accounting for satellites. %Because of the large systematics that are common in contemporary group-finding methods \citep{campbell_etal15}, we consider forward modeling methods to be a requisite for continued progress on this topic.

@@

\ben
\item Small galaxies cluster more strongly than large galaxies of the same stellar mass. Differences between the clustering of small and large galaxies increase on small scales $R\lesssim1\mpc,$ and decrease with stellar mass.
\item The most important ingredient influencing this signal is the relative size of central and satellite galaxies. The magnitude, scale-dependence, and $\mstar-$dependence of $\rhalf-$dependent clustering provides strong evidence that satellite galaxies are smaller than central galaxies of the same halo mass.
\item A simple empirical model in which $\rhalf$ is set by halo $\rvir$ at the time of peak halo mass exhibits a clustering signal that is strikingly similar to that r seen in SDSS.
\item Models in which $\rhalf$ is regulated by $\mstar,$ rather than $\mhalo,$ are grossly discrepant with the observed clustering signal, even when accounting for satellite mass stripping.
\item Taken together, our findings indicate that satellite-specific processes play a sub-dominant role in setting the relative size of centrals and satellites, which instead appears to be largely predetermined at the time of satellite infall.
\een

Our results can be treated as a starting point for more complex and fine-grained models of galaxy size, such as semi-analytic models and hydrodynamical simulations. For convenience, Figure \ref{fig:censatsizeratios} provides a simple summary statistic that can be used as a diagnostic for calibrating alternative modeling efforts. We view the present work as a pilot study that motivates a Bayesian inference program to tightly constrain the galaxy size-halo connection with forward modeling techniques, in direct analogy to the literature on the stellar-to-halo-mass relation. Our publicly available python code provides a simple means for cosmological surveys to generate synthetic galaxy populations with realistic sizes across the cosmic web.

@@

We thank the {\tt Astropy} developers for the package-template \citep{astropy}, and {\tt NumPy} \citep{numpy_ndarray}, {\tt SciPy} \citep{scipy}, IPython, Matplotlib, and GitHub for their extremely useful free software. While writing this paper we made extensive use of the Astrophysics Data Service (ADS) and {\tt arXiv} preprint repository.

This research was supported in part by the National Science Foundation under Grant No. NSF PHY11-25915. Work done at Argonne National Laboratory was supported under the DOE contract DE-AC02-06CH11357. AK was  supported by NSF grant  AST-1412107 and by the Kavli Institute for Cosmological Physics at the University of Chicago through grant PHY-1125897 and an endowment from the Kavli Foundation and its founder Fred Kavli.
